All Samples Model

### 2018-03-06 AllSamplesModel.Rmd
adding to pipeline

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2 October 2017"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

@@

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')

data = '~/Data/Ellie'
info.dir = '~/Data/Ellie/QDNAseq'
#tile.w=c(5e06, 'arms')
#tile.w=5e+06
#tile.w=1e7
#tile.w='arms'

#includeDemo = F

#data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
#analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

## Hospital.Research.ID info file

patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

if (length(patient.file) != 1 | length(demo.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", info.dir))

@@

patient.info = subset(all.patient.info, Set == 'Training')

validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.validation.info = summarise.patient.info(validation.patient.info)

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

@@

pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')

all.info = rbind(sum.patient.data, sum.validation.info)

@@

```{r echo=F}

pander(
  as.data.frame(all.info %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range'=paste(range(end.year-start.year), collapse='-'),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
))
, caption='full cohort')
```

```{r echo=F}
cohortSummary = as.data.frame(all.info %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
))

cohortSummary = cbind(cohortSummary,
  rbind(table(subset(all.patient.info, Status == 'NP')$Pathology),
        table(subset(all.patient.info, Status == 'P')$Pathology)))

  rbind(table(subset(patient.info, Status == 'NP')$Pathology),
        table(subset(patient.info, Status == 'P')$Pathology))

  rbind(table(subset(validation.patient.info, Status == 'NP')$Pathology),
        table(subset(validation.patient.info, Status == 'P')$Pathology))

@@

`r pander(t(cohortSummary), caption='Full cohort, split by status')`

@@

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=F)
dim(dysplasia.df)
length(labels)

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}

chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

@@

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples') + theme(text=element_text(size=12))
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)

gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)

@@

```{r loaddata, message=F, warning=F, echo=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

@@

ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)

@@

```{r echo=F, fig.height=5, fig.width=5}
library(ggfortify)
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme

ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
```

@@

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

@@

ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

@@

p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')

@@

rm(plots, performance.at.1se,coefs, nzcoefs, fits)
```

ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)

@@

ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

p = globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

```{r, eval=F, echo=F, message=F}
image(cov(dysplasia.df[names(which(labels == 1)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

image(cor(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))], use='p'), breaks=seq(-1,1,l=21), col=colorRampPalette(c('red','black','green'))(20),  useRaster = T)

chr <- as.numeric(sub(":.+","",colnames(dysplasia.df[names(which(labels == 0)),c(1:(ncol(dysplasia.df)-3))])))
c <- which(diff(chr)==1)/length(chr)
abline(h=c, v=c, col='blue')

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Samplename','Prediction')]))
rownames(preds) = preds$Samplename
p <- preds[names(labels), 'Prediction']

d <- dist(dysplasia.df[,c(1:(ncol(dysplasia.df)-3))])
t <- tsne::tsne(d)
plot(t,  bg=colorRampPalette(c('green','grey','red'))(10)[cut(p, breaks = seq(0,1,l=11))], pch=21, col=ifelse(labels==1,"black","white"))
```

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'
roc$is = 'discovery'

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

insilico.rocs[['discovery']] = roc

@@

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, is == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
thresholds = 
  rbind.data.frame(
#    est=t(round(pROC::coords(roc, "best"),2)),
    max=t(round(pROC::coords(roc, 0.8),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6),2))
)
thresholds$label=rownames(thresholds)

# ggplot() + geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() + 
#   geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
#   geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
#     paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(3,'Dark2')) + scale_color_manual(values=brewer.pal(3,'Dark2')) +
#   labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

roc.plot(rocNOHGD, 'Model excluding HGD/IMC')
pROC::coords(rocNOHGD, "best")
pROC::coords(rocNOHGD, 0.6)
pROC::auc(rocNOHGD, partial.auc=c(0.89,0.9), partial.auc.focus='spec', partial.auc.correct=T)

@@

```{r pathroc, echo=F, fig.height=4, fig.width=8 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'
ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
rocPtM$model = 'Per Patient Max'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
#roc.plot(rocPtA)
rocPtA$model = 'Per Patient Avg'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

# get.auc.at(p53roc)
# get.auc.at(hgdpathroc)
# get.auc.at(pathroc)
# get.auc.at(roc)

roc$model = "CN Model"

m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)

rocList = list(roc,pathroc,hgdpathroc,p53roc)

@@

ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

@@

ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)

#multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', palette='Dark2')
```

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR', 'Sample.Mean','Sample.SD','Sample.Var')] %>% group_by(PID, Samplename) %>% mutate(
  'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

@@

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status', 'Total.Reads')])
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD','IMC'), ordered = T)

@@

back$Pathology = as.character(back$Pathology)
#back$Pathology[which(back$Pathology %in% c('HGD','IMC'))] = 'IMC/HGD'
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)

ggplot(back, aes(Pathology, Prediction, group=Pathology, fill=Pathology)) + facet_grid(~Status) + 
  geom_jitter(width=0.2) + geom_boxplot() + plot.theme  + theme(legend.position = 'none')

@@

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 'se'=(sd(Prediction)/sqrt(length(Prediction)))*1.96 ))
#ggplot(qt,aes(quants, mn, ymin=mn-se, ymax=mn+se)) + geom_point() + geom_errorbar() + labs(x='Deciles', y='Mean probability', title='CI around probabilities by decile')

pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)
```

>>>>

---
title: "All Samples Model"
author: "Sarah Killcoyne"
date: "16 Feb 2018"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

@@

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/Ellie'

cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
#analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

@@

patient.info = arrange(all.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

@@

@@

```{r echo=F}
pander(
  as.data.frame(sum.patient.data %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range'=paste(range(end.year-start.year), collapse='-'),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
)), caption='full cohort', justify='left')
```

```{r echo=F}
cohortSummary = as.data.frame(sum.patient.data %>% group_by(Status) %>% summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
))

cohortSummary = cbind(cohortSummary,
  rbind(table(subset(all.patient.info, Status == 'NP')$Pathology),
        table(subset(all.patient.info, Status == 'P')$Pathology)))
  rbind(table(subset(patient.info, Status == 'NP')$Pathology),
        table(subset(patient.info, Status == 'P')$Pathology))

@@

`r pander(t(cohortSummary), caption='Full cohort, split by status', justify='left')`

@@

```{r labelsPNP, echo=F, message=F, include=T}
file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=T)
dim(dysplasia.df)
length(labels)

@@

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

@@

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples') + theme(text=element_text(size=12))
#ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
gp

gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") + theme(text=element_text(size=12))
gbe
#ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)

@@

```{r loaddata, message=F, warning=F, echo=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

@@

#ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

#ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

@@

#ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)

@@

@@

#ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}
nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
#ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

@@

#ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

@@

#ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)

@@

p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
  #geom_text_repel(aes(label=pt)) +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p

#ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'No HGD')

@@

rm(plots, nzcoefs, fits)
```

@@

#ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

#ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

@@

p = globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
p

#ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

#ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

@@

#ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

@@

@@

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'
roc$is = 'all'

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
#ggsave(filename='plots/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

insilico.rocs[['all']] = roc

@@

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, is == 'all'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
#ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

@@

```{r pathroc, echo=F, fig.height=4, fig.width=10 }
pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'
#ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
#ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
#ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
#ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
rocLastEndo$model = 'Last endoscopy'
#ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
rocFirstEndo$model = 'First endoscopy'
#ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
rocPtM$model = 'Per Patient Max'
#ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
#roc.plot(rocPtA)
rocPtA$model = 'Per Patient Avg'
#ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

@@

roc$model = "CN Model"
m = rbind.data.frame(
  #get.roc.stat(p53roc),
  get.roc.stat(hgdpathroc),
  get.roc.stat(pathroc),
  get.roc.stat(roc))
m$model = factor(m$model, levels=c('HGD Only','HGD & LGD','CN Model'), ordered = T)

rocList = list(roc,pathroc,hgdpathroc)
multi.roc.plot(rocList)

@@

#ggsave('plots/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

@@

#ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)

multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', colors=rev(cols[1:7]))
```

@@

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% mutate(
    'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

@@

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status')])

@@

back$Pathology = as.character(back$Pathology)
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)

@@

qt = cbind(qt, back %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

#ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

mod = subset(back, Risk == 'Moderate')

```{r relrisk, echo=F, fig.width=6, fig.height=6}
myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR

#ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
predhist

#ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
